---
title: "HW3"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(RColorBrewer)
artist <- read.csv("artistdata.csv")
```
Turn in the knitted pdf of an R Markdown document that shows the distribution of birth decade for various races/ethicities of artists, based on the data available in the artist data file I shared. Produce at least two different graphs to visualize this information.

\vspace{5mm}

The original dataset was subset to remove all rows that had values of "NA" for either artist ethnicity or birth year.
```{r}
artist<- artist[complete.cases(artist[,c("ethnicity", "year")]),]
```

#Density curves for artist ethnicity
The plot below shows the density for artist birth year for the different races and ethnicities. As can be seen, the majority of artists identified as black, hispanic, and other were born after 1800, although hispanic artists were also present between 1600 and 1750.
```{r}
ggplot(data = artist, aes(x=year)) + 
  geom_density(aes(color = ethnicity), size=1, adjust=1.5, alpha=0.5) +
  scale_color_brewer(palette="Dark2") +
  ggtitle(label="Density of Artist Ethnicity By Birth Decade") + 
  coord_cartesian(xlim = c(1500,2000)) +
  scale_y_continuous(expand =c(0, 0), limits = c(0,.018), breaks = seq(0,.018,.002)) +
  theme_minimal()
```

\newpage

#Histograms for artist ethnicities
The histograms below show the distribution of artists for their birth years, separated by race/ethnicity. Each graph has a different y-axis scale due to the difference in total artists by each race. As there were only 75 artists in the largest bin for those who were identified as other, this distribution was difficult to see when using the same y-axis scale as for artists identified as white, as there are over 2,000 in the largest bin for their distribution.
```{r}
ggplot(artist, aes(x=year, fill=ethnicity)) + 
  geom_histogram(binwidth=50) +
  scale_fill_brewer(palette="Dark2") +
  coord_cartesian(xlim=c(1500,2000)) +
  facet_wrap(~ethnicity, scales = "free") +
  scale_y_continuous(breaks = scales::pretty_breaks(10), limits = c(0, NA)) +
  guides(fill=FALSE) +
  theme_minimal()
```

